Dismal Code: Studying the Evolution of Security Bugs

### 2013-04-08 docs/evol.tex
correct bib

A security bug is a programming error that introduces a potentially exploitable
weakness into a computer system~\cite{SSL12}\cite{TJBD11}\cite{M06}. This
weakness could lead to a security breach with unfortunate consequences in
different layers, like databases, native code, applications, libraries and
others. Despite the significant effort to detect and eliminate such
bugs~\cite{SZ12}, little attention has been paid to study them in relation to
software evolution~\cite{L96}\cite{LRWPT97}\cite{IB06}\cite{RGMA06}. 
One of the most common approaches to identify security bugs is
{\it static analysis}~\cite{CW07}. This kind of analysis involves the
inspection of the program's source or object code without executing
it.

In this
paper we present how we used a large software ecosystem to analyse how different
types of security vulnerabilities are related to the evolving software packages.
For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10}\cite{HP07}\cite{HP04}\cite{HW08}\cite{SHP06}.
Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the
{\it Maven Central Repository}\footnote{\url{http://search.maven.org/}}
(approximately 265GB of data---see Section~\ref{sec:data}).
Then we observed the changes that involved the security bugs and their characteristics.
This research builds upon our earlier work on the topic~\cite{MGS12}.

We chose to focus our study on security bugs rather than other types of
software defects. This is because compared to other bug categories,
failures due to security bugs have two distinct features:
they can severeyl affect an organization's infrastructure\cite{SZ12}, and
they can cause significant financial damage to
an organization~\cite{TH04}\cite{BCL08}\cite{R06}. 
Specifically,
whereas a software bug can cause a software artifact to fail,
a security bug can allow a malicious user to alter the execution
of the entire application for his or her own gain.
In this case, such bugs could give rise to a wide
range of security and privacy issues, like
the access of sensitive information,
the destruction or modification of data, and
denial of service.
Moreover, security bug disclosures lead to a negative and significant change
in market value for a software vendor~\cite{TW07}.
Hence, one of the basic pursuits in every new software release should
be to mitigate such bugs.

% @dimitro: say why we did this work, not how we did it
The motivation behind our work was to validate whether programmers care for
the risk posed by security bugs when they release a new version of their software.
In addition, we wanted to investigate other critical features associated with such
vulnerabilities like the persistence of a bug;
in essence, to see whether critical bugs stay unresolved for a long time.
Also, we wanted to elaborate more on the relation of security
bugs with other bug categories.
In the same manner, we tried to looked at the relationship
between successive product versions and the number of security bugs,
knowing the that research has produced contradictory results on this
issue~\cite{BP84}\cite{SYTP85}\cite{NBZ06}\cite{GKMS00}.
Finally, we examined the Maven ecosystem as a whole from a security
perspective. Its structure gave us the opportunity to see if a project version that is a dependency to
a large number of others contains a low rate of security bugs~\cite{MW10}.
%In this case we can strengthen the relationship between Linus' Law and
%software security as Meneely et al. have already indicated~\cite{MW10}.

@@

{\it Querying techniques} are used to answer a broad range of questions
regarding the evolution history of a project~\cite{HG05}. Bhattacharya et
al.~\cite{BN11}\cite{B11} proposed a framework that is based upon
recursively enumerable languages. The framework can correlate software
bugs with developers in various ways. For instance, return the list of
bugs fixed by a specific developer. Fischer et al.~\cite{FPG03} proposed
an approach for populating a release history database that combines code
information with bug tracking data. In this way, a developer can couple files
that contain common bugs, estimate code maturity with respect to the bugs
and others. The ``Ultimate Debian Database''~\cite{NZ10} is an {\sc sql}-based
framework that integrates information about the Debian project from various
sources to answer queries related to software bugs and source code.

@@

The evolution of software artifacts has also been analysed to {\it reduce the false
alarms} of the various static analysis tools. To achieve this, Spacco et
al.~\cite{SHP06} introduced {\it pairing} and {\it warning signatures}. In the
former, they tried to pair sets of bugs between versions in order to find
similar patterns. In the latter, they computed a signature for every bug. This
signature contained elements like the name of the class where the bug was found,
the method and others. Then they searched for similar signatures between
versions. In their research they also studied the evolution of 116 sequential
builds of the Sun Java Sevelopment Kit ({\sc jdk}). Their findings indicated that
high priority bugs are fixed over time. To improve the precision of bug
detection, Kim et al.~\cite{KE07b}\cite{KE07} proposed a history-based warning
prioritization algorithm by mining the history of bug-fixes of three
different projects. The algorithm was based on the following hypothesis: if a
bug that belongs to a specific category is eliminated by a code change, then
this warning category is important. Working towards the same direction, Heckman
et al.~\cite{HW09}\cite{HW08} have introduced benchmarks that use specific
correlation algorithms and classification techniques to evaluate alert
prioritization approaches.

>>>>

A security bug is a programming error that introduces a potentially exploitable
weakness into a computer system~\cite{SSL12,TJBD11,M06}. This
weakness could lead to a security breach with unfortunate consequences in
different layers, like databases, native code, applications, libraries and
others. Despite the significant effort to detect and eliminate such
bugs~\cite{SZ12}, little attention has been paid to study them in relation to
software evolution~\cite{L96,LRWPT97,IB06,RGMA06}. 
One of the most common approaches to identify security bugs is
{\it static analysis}~\cite{CW07}. This kind of analysis involves the
inspection of the program's source or object code without executing
it.

In this
paper we present how we used a large software ecosystem to analyse how different
types of security vulnerabilities are related to the evolving software packages.
For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10,HP07,HP04,HW08,SHP06}.
Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the
{\it Maven Central Repository}\footnote{\url{http://search.maven.org/}}
(approximately 265GB of data---see Section~\ref{sec:data}).
Then we observed the changes that involved the security bugs and their characteristics.
This research builds upon our earlier work on the topic~\cite{MGS12}.

We chose to focus our study on security bugs rather than other types of
software defects. This is because compared to other bug categories,
failures due to security bugs have two distinct features:
they can severeyl affect an organization's infrastructure\cite{SZ12}, and
they can cause significant financial damage to
an organization~\cite{TH04,BCL08,R06}. 
Specifically,
whereas a software bug can cause a software artifact to fail,
a security bug can allow a malicious user to alter the execution
of the entire application for his or her own gain.
In this case, such bugs could give rise to a wide
range of security and privacy issues, like
the access of sensitive information,
the destruction or modification of data, and
denial of service.
Moreover, security bug disclosures lead to a negative and significant change
in market value for a software vendor~\cite{TW07}.
Hence, one of the basic pursuits in every new software release should
be to mitigate such bugs.

% @dimitro: say why we did this work, not how we did it
The motivation behind our work was to validate whether programmers care for
the risk posed by security bugs when they release a new version of their software.
In addition, we wanted to investigate other critical features associated with such
vulnerabilities like the persistence of a bug;
in essence, to see whether critical bugs stay unresolved for a long time.
Also, we wanted to elaborate more on the relation of security
bugs with other bug categories.
In the same manner, we tried to looked at the relationship
between successive product versions and the number of security bugs,
knowing the that research has produced contradictory results on this
issue~\cite{BP84,SYTP85,NBZ06,GKMS00}.
Finally, we examined the Maven ecosystem as a whole from a security
perspective. Its structure gave us the opportunity to see if a project version that is a dependency to
a large number of others contains a low rate of security bugs~\cite{MW10}.
%In this case we can strengthen the relationship between Linus' Law and
%software security as Meneely et al. have already indicated~\cite{MW10}.

@@

{\it Querying techniques} are used to answer a broad range of questions
regarding the evolution history of a project~\cite{HG05}. Bhattacharya et
al.~\cite{BN11,B11} proposed a framework that is based upon
recursively enumerable languages. The framework can correlate software
bugs with developers in various ways. For instance, return the list of
bugs fixed by a specific developer. Fischer et al.~\cite{FPG03} proposed
an approach for populating a release history database that combines code
information with bug tracking data. In this way, a developer can couple files
that contain common bugs, estimate code maturity with respect to the bugs
and others. The ``Ultimate Debian Database''~\cite{NZ10} is an {\sc sql}-based
framework that integrates information about the Debian project from various
sources to answer queries related to software bugs and source code.

@@

The evolution of software artifacts has also been analysed to {\it reduce the false
alarms} of the various static analysis tools. To achieve this, Spacco et
al.~\cite{SHP06} introduced {\it pairing} and {\it warning signatures}. In the
former, they tried to pair sets of bugs between versions in order to find
similar patterns. In the latter, they computed a signature for every bug. This
signature contained elements like the name of the class where the bug was found,
the method and others. Then they searched for similar signatures between
versions. In their research they also studied the evolution of 116 sequential
builds of the Sun Java Sevelopment Kit ({\sc jdk}). Their findings indicated that
high priority bugs are fixed over time. To improve the precision of bug
detection, Kim et al.~\cite{KE07b,KE07} proposed a history-based warning
prioritization algorithm by mining the history of bug-fixes of three
different projects. The algorithm was based on the following hypothesis: if a
bug that belongs to a specific category is eliminated by a code change, then
this warning category is important. Working towards the same direction, Heckman
et al.~\cite{HW09,HW08} have introduced benchmarks that use specific
correlation algorithms and classification techniques to evaluate alert
prioritization approaches.
